### 2021-05-20 presentation.Rmd
converted presentation

---
title: "Cool Title"
author: "Chris Tastad"
date: "`r format(Sys.time(), '%b %d, %Y')`"
output:
  revealjs::revealjs_presentation:
    center: false
    theme: night
    highlight: zenburn
    transition: none
    reveal_options:
      controls: false
      height: "100%"
      margin: 0.04
      slideNumber: true
      previewLinks: true
---

## Technical Debt fig

## CCGD Overview {.center}

## Breakdown of build table process

```{r global_options, include = F}
knitr::opts_chunk$set(eval = F, echo = F, warning = F, message = F)
```

## Application Concept

## Forward Transposon Screens

## Project Requirements

::: {#table:projectReqs}
      Feature           Goal                       Framework   Description
  --- ----------------- -------------------------- ----------- -------------------------------------------------------------------------------------------------------------------------------------------------------------------------
  1   server OS         upgrade server OS          RHEL7       Transition of architecture for public server host. This is required by University OIT due to end-of-life schedule for RHEL6.
  2   web front-end     rewrite web interface      Rmarkdown   Improvements to the web interface written in a modern, simplified language. This improved access to content creation and allowed for automation in front-end rendering.
  3   table build       rewrite table build        Rshiny      Rshiny offered a dramatic improvement to replace the existing process by merging the table build back-end with a modern web display of the app interface.
  4   content update    improved admin controls    BASH/R      Old app version confined some content controls to app author, limiting ability to make contributions by product owner (Tim Starr).
  5   version control   implement best practices   git/docs    No version control was used in the original development of the app. This and other documentation practices were expanded in the rewrite.
  6   multiple          resource improvements      codebase    Due to the lack of some best practices, there were many opportunities to make impactful resource improvements.

:::

## Rmarkdown Rendering Flow
<img src="./fig/rmarkdown_flow.svg">

## YAML Configuration Segments

<div class="column">

```{r echo = T}
title: "Candidate Cancer Gene Database"
bibliography: refs/ccgd_paper.bib
nocite: '@*'
output:
    html_document:
        includes:
            in_header: "styles/favicon.html"
            after_body: "styles/footer.html"
        css: styles/styles.css
        theme: readable
---

```

```{r echo = T}
name: "Candidate Cancer Gene Database"
navbar:
    title: "CCGD"
    left:
        - text: "Home"
          href: index.html
        - text: "Search"
          href: search.html
        - text: "Help"
          href: help.html
        - text: "References"
          href: references.html
        - text: "Contact"
          href: contact.html

```

</div>

![](./fig/yaml_illustration.svg)

## Subroutine Flow
![](./fig/subroutine_flow.svg)

## Table Reactive Graph
![](./fig/reactive_graph.svg)

## Server-side Processing Flow
![](./fig/parse_flow.svg)

## Search Interface Changes
![](./fig/interface.svg)

## Disk Use Comparison
![](./fig/space_tree.svg)

## Project Directory File Tree Comparison
![](./fig/file_tree.svg)

## Documentation

## app.R code

## markdown code

## List of app improvements

## The End {.center}

<!-- style -->

<style>

body {
  background: #2e3440;
}

.column {
  column-count: 2;
}

div.sourceCode {
  background-color: #2e3440;
}

.sourceCode {
  background-color: #393f4b;
}

.reveal .progress span {
  background: #7f9f7f;
}

.reveal section img{
  border: 30px solid #e0e1e2;
  border-radius: 10px;
  }

</style>

<!-- slide templates

## Section Title {.center}

## Multi-Column Slide

<div class="column">

### column1
- some cool words
- https://www.merriam-webster.com/dictionary/wherever

### column2
- some more cool words

</div>

## Figure Slide
```{r eval=T}
attach(mtcars)
par(mfrow=c(2,2))
plot(wt,mpg, main="Scatterplot of wt vs. mpg")
plot(wt,disp, main="Scatterplot of wt vs disp")
hist(wt, main="Histogram of wt")
boxplot(wt, main="Boxplot of wt")
```

## Code Slide

```{r echo = T}
x = 1
if x == 1:
    # indented four spaces
    print("x is 1.")
```

-->